---
title: "Initial Analyses"
output: html_notebook
---

```{r echo=F}
library(tidyverse); library(reshape2); library(forcats)
library(psych); library(lme4); library(TripleR)
options(tibble.width = Inf)


setwd('~/Dropbox/Twitter and personality project/personality impressions study/Personality Ratings Study/third times a charm')

df <- tbl_df(read_csv('Personality Impressions on Twitter_Filtered_052317.csv'))
```

## Restructure the Data

The current dataset is very long -- each row is a single response to a single question. To better facilitate some of the upcoming analyses, we'll restructure into a wider format. Here, each row will contain all of the responses from one participant about one target. Since each participant rated a set of 5 targets as well as themselves (for demographic items), each participant will have a total of 6 rows for each kind of stimuli (note that participants who completed several different surveys (e.g., profiles and followers) are still included, so they can have up to 18 rows).

Make the dataset wide, dropping the variables that identify the question numbers (they're redundant with the labels but less informative):
```{r}
df_wide<-dcast(df %>% select(-c(question, orig_quest_num)),
                pID + target_user + stim + study + ...				
               ~qlabel,  					
               value.var="value")

df_wide %>% arrange(pID, target_user) %>% head()
```



## Scoring Big 5 Scales

Before looking too closely, we'll start by reverse-coding and scoring the B5 scales (from BFI-10). Note that we did ask a handful of extraversion-related questions, but will only score the two that come from the 10-item version of the scale.

```{r}
df_wide <- df_wide %>% 
        mutate(cold_r = 6-cold, 
                   reserved_r = 6 - reserved,
                   lazy_r = 6 - lazy,
                   relaxed_r = 6 - relaxed,
                   fewartint_r = 6 - fewartint) %>%
        mutate(ext = (outgoing + cold_r)/2,
               agr = (trusting+reserved_r)/2,
               con = (thorough + lazy_r)/2,
               neu = (nervous + relaxed_r)/2,
               ope = (actimag + fewartint_r)/2)
        

```

Convert back to long:

```{r}
df2<- df_wide %>% gather(qlabel, # name of the new key variable
                         value, # name of the new value variable
                         c(actimag:introduceFriend, lazy:ope), # variables to make long; all of them except know
                         na.rm = T) %>% # drop rows where value is missing
        arrange(pID, target_user)

(df2)

```

## Pre-Processing Factor Vars  
 
Now that we're in wide format, it'll be easier to change some of the less intuitive ways Qualtrics encoded some of the values. We'll also re-code the categorical variables to give them more meaningful values, and transform some of our character vars to factors so they're easier to work with. Note that all of the likert responses are 1 (strongly disagree) to 5 (strongly agree), with the exception of SES, which is 1 (low) to 10 (high).  

Convert to factors:

```{r}
df_wide <- df_wide %>% mutate(sex = factor(sex),
                              sub_sex = factor(sub_sex),
                              sub_sexOrientation = factor(sub_sexOrientation),
                              race = factor(race),
                              sub_EngFirstLang = factor(sub_EngFirstLang),
                              sub_durationEng = factor(sub_durationEng),
                              sub_socMediaUse = factor(sub_socMediaUse),
                              sub_twitteruse = factor(sub_twitteruse),
                              sample = factor(sample),
                              pID = factor(pID),
                              target_user = factor(target_user),
                              stim = factor(stim),
                              study = factor(study),
                              sub_raceAsian = factor(sub_raceAsian),
                              sub_raceBlk = factor(sub_raceBlk),
                              sub_raceHisp = factor(sub_raceHisp),
                              sub_raceNatAm = factor(sub_raceNatAm),
                              sub_raceOther = factor(sub_raceOther),
                              sub_raceWht = factor(sub_raceWht),
                              know = factor(know)
                              )
```

```{r}
summary((df_wide$sex))
summary((df_wide$sub_sex))
summary((df_wide$sub_sexOrientation))
```
```{r}
summary((df_wide$race))
```

```{r}
summary((df_wide$sub_EngFirstLang))
summary((df_wide$sub_durationEng))
```

```{r}
summary((df_wide$sub_socMediaUse))
summary((df_wide$sub_twitteruse))
```



```{r}
df_wide <- df_wide %>% mutate(sex = fct_recode(sex, 
                                    "male" = "1",
                                    "female" = "2",
                                    "other" = "3"),
                   sub_sex = fct_recode(sub_sex, 
                                    "male" = "1",
                                    "female" = "2",
                                    "other" = "3"),
                   sub_sexOrientation = fct_recode(sub_sexOrientation,
                                                   "homosexual" = "1",
                                                   "heterosexual" = "2",
                                                   "bisexual" = "3",
                                                   "other" = "4"),
                   race = fct_recode(race, 
                                    "black" = "1",
                                    "asian" = "2",
                                    "hispanic" = "3",
                                    "nativeAmerican" = "4",
                                    "nativeHIpacisl" = "5",
                                    "white" = "6",
                                    "other" = "7"),
                   sub_EngFirstLang = fct_recode(sub_EngFirstLang, 
                                    "yes" = "1",
                                    "no" = "2"),
                   sub_durationEng = fct_recode(sub_durationEng, 
                                    "0-6 months" = "1",
                                    "7-12 months" = "2",
                                    "1-2 years" = "3",
                                    "3-4 years" = "4",
                                    "5-6 years" = "5",
                                    "more than 6 years" = "6"),
                   sub_socMediaUse = fct_recode(sub_socMediaUse, 
                                    "never" = "1",
                                    "less than once per month" = "2",
                                    "once per month" = "3",
                                    "2-3x per month" = "4",
                                    "once per week" = "5",
                                    "2-3x per week" = "6",
                                    "daily" = "7"),
                   sub_twitteruse = fct_recode(sub_twitteruse, 
                                    "never" = "1",
                                    "less than once per month" = "2",
                                    "once per month" = "3",
                                    "2-3x per month" = "4",
                                    "once per week" = "5",
                                    "2-3x per week" = "6",
                                    "daily" = "7"),
                   know = fct_recode(know, "yes" = "1", "no" = "2"))

```

```{r}
summary((df_wide$sex))
summary((df_wide$sub_sex))
summary((df_wide$sub_sexOrientation))
```
```{r}
summary((df_wide$race))
```

```{r}
summary((df_wide$sub_EngFirstLang))
summary((df_wide$sub_durationEng))
```

```{r}
summary((df_wide$sub_socMediaUse))
summary((df_wide$sub_twitteruse))
```




## Descriptives

Let's take a look at descriptives by each survey.

Study 1:
```{r}
s1p <- suppressWarnings(data.frame(describe(df_wide %>% 
                                                    filter(study == 1 
                                                           & stim == 'profile'))[19:93,1:4]))

s1p <- s1p %>% mutate(vars = rownames(s1p)) %>% filter(n > 0) %>% 
        rename(N_profile = n, M_profile = mean, SD_profile = sd)

s1fo <- suppressWarnings(data.frame(describe(df_wide %>% 
                                                     filter(study == 1 
                                                            & stim == 'followers'))[19:93,1:4]))

s1fo <- s1fo %>% mutate(vars = rownames(s1fo)) %>% filter(n > 0)%>% 
        rename(N_followers = n, M_followers = mean, SD_followers = sd)

s1f <- suppressWarnings(data.frame(describe(df_wide %>% 
                                                    filter(study == 1 
                                                           & stim == 'friends'))[19:93,1:4]))

s1f <- s1f %>% mutate(vars = rownames(s1f)) %>% filter(n > 0)%>% 
        rename(N_friends = n, M_friends = mean, SD_friends = sd)


merge(s1p,s1f, by = 'vars') %>% merge(s1fo, by = 'vars')

```


Study 2:  
```{r}
s2p <- suppressWarnings(data.frame(describe(df_wide %>% 
                                                    filter(study == 2))[19:93,1:4]))

s2p %>% mutate(vars = rownames(s2p)) %>% filter(n > 0)

```

```{r}

s1q_lab <-c('outgoing','cold','thorough','nervous','actimag','reserved',
            'trusting','lazy','relaxed','fewartint','trustworthy','selfesteem',
            'funny','assertive','arrogant','intelligent','impulsive',
            'centerattn', 'attractive','ilike')



ggplot(df2 %>% filter(qlabel %in% s1q_lab), aes(value)) + facet_wrap(~qlabel) + 
        geom_histogram(data = df %>% filter(qlabel %in% s1q_lab & stim =='profile'), 
                       fill = 'red', alpha = .4, bins = 10) +
        geom_histogram(data = df %>% filter(qlabel %in% s1q_lab & stim =='friends'), 
                       fill = 'blue', alpha = .5, bins = 10) +
        geom_histogram(data = df %>% filter(qlabel %in% s1q_lab & stim =='followers'), 
                       fill = 'green', alpha = .2, bins = 10) 

```


```{r}
ggplot(df %>% filter(qlabel %in% s1q_lab), aes(value, colour = stim, fill = stim)) + 
        facet_wrap(~qlabel) +
        geom_density(alpha = .1)
```




```{r}
s1q_targdemo <- c('age', 'sex', 'ses','race','know')

ggplot(df %>% filter(qlabel %in% s1q_targdemo), aes(value)) + 
        facet_wrap(~qlabel, scales = 'free_x') + 
        geom_histogram(data = df %>% filter(qlabel %in% s1q_targdemo & stim =='profile'), 
                       fill = 'red', alpha = .4, bins = 10) +
        geom_histogram(data = df %>% filter(qlabel %in% s1q_targdemo & stim =='friends'), 
                       fill = 'blue', alpha = .5, bins = 10) +
        geom_histogram(data = df %>% filter(qlabel %in% s1q_targdemo & stim =='followers'), 
                       fill = 'green', alpha = .2, bins = 10) 
```

```{r}
s1q_targdemo2 <- c('age', 'sex', 'ses','race')
ggplot(df %>% filter(qlabel %in% s1q_targdemo2), aes(value, colour = stim, fill = stim)) + 
        facet_wrap(~qlabel, scales = 'free_x') +
        geom_density(alpha = .1)

```



```{r}
s2q_lab <- c('trustOrganize','followSM','helpConflict','hire', 'friendshipOnline',
             'helpIntellChallenge','movieTVrec','introduceFriend','adviceGoodImpression',
             'romanticPartner_self','triviaPartner','casualSocial','trustInfoOnline',
             'groupProject','buySomethingOnline','interestNewsOnline','askNotes',
             'romanticPartner_friend','converseIntellStim','workTaskConflict',
             'valueOpinion','trustDeadlines','friendsOffline','workStressfulTask',
             'musicRec','paperFeedback')

ggplot(df %>% filter(qlabel %in% s2q_lab), aes(value)) + facet_wrap(~qlabel) + 
        geom_histogram(bins = 10)
```

# ICCs

This will generate ICCs for all of the perception variables, with no additional filtering.

```{r}
skip <- c('sub_age','sub_durationEng','sub_EngFirstLang', 'sub_raceAsian',
          'sub_raceBlk','sub_raceHisp','sub_raceNatAm','sub_raceOther',
          'sub_raceWht','sub_SES','sub_sex','sub_sexOrientation',
          'sub_socMediaUse','sub_twitteruse')

ICCs<-df2 %>% filter(!(qlabel %in% skip)) %>%
        group_by(study, stim, qlabel) %>%
        do({
                model<-lmer(value~1 + (1|pID) + (1|target_user), data=.)
                var<-as.data.frame(VarCorr(model))
                ICC.p<-round((var[which(var$grp=="pID"),4]/sum(var[,4])), 
                             digits = 3)
                ICC.t<-round((var[which(var$grp=="target_user"),4]/sum(var[,4])), 
                             digits = 3)
                data.frame(ICC.t,ICC.p)
        })
```

Filter out cases where the participant said they knew the target.
```{r}
ICCs_unknown<-df2 %>% filter(know == 2) %>%
        filter(!(qlabel %in% skip)) %>% 
        group_by(study, stim, qlabel) %>%
        do({
                model<-lmer(value~1 + (1|pID) + (1|target_user), data=.)
                var<-as.data.frame(VarCorr(model))
                ICC.p_unknown<-round((var[which(var$grp=="pID"),4]/sum(var[,4])), 
                             digits = 3)
                ICC.t_unknown<-round((var[which(var$grp=="target_user"),4]/sum(var[,4])), 
                             digits = 3)
                data.frame(ICC.t_unknown,ICC.p_unknown)
        })
```
```{r}
iccs <- merge(ICCs,ICCs_unknown, on = c('study','stim','qlabel'))
```


Filter out all subsequent attempts (take just the first attempt overall, instead of just the 1st attempt for a particular stimulus).

```{r}
ICCs_1st<-df2 %>% filter(attempt==1) %>%
        filter(!(qlabel %in% skip)) %>% 
        group_by(study, stim, qlabel) %>%
        do({
                model<-lmer(value~1 + (1|pID) + (1|target_user), data=.)
                var<-as.data.frame(VarCorr(model))
                ICC.p_1st<-round((var[which(var$grp=="pID"),4]/sum(var[,4])), 
                             digits = 3)
                ICC.t_1st<-round((var[which(var$grp=="target_user"),4]/sum(var[,4])), 
                             digits = 3)
                data.frame(ICC.t_1st,ICC.p_1st)
        })
```
```{r}
iccs <- merge(iccs, ICCs_1st, on = c('study','stim','qlabel'))
```


Filter out all subsequent attempts AND knowns.

```{r}
ICCs_1stunknown<-df2 %>% filter(know == 2 & attempt == 1) %>%
        filter(!(qlabel %in% skip)) %>% 
        group_by(study, stim, qlabel) %>%
        do({
                model<-lmer(value~1 + (1|pID) + (1|target_user), data=.)
                var<-as.data.frame(VarCorr(model))
                ICC.p_1stunknown<-round((var[which(var$grp=="pID"),4]/sum(var[,4])), 
                             digits = 3)
                ICC.t_1stunknown<-round((var[which(var$grp=="target_user"),4]/sum(var[,4])),
                             digits = 3)
                data.frame(ICC.t_1stunknown,ICC.p_1stunknown)
        })
```

```{r}
iccs <- merge(iccs, ICCs_1stunknown, on = c('study','stim','qlabel'))
head(iccs)
```

### test differences in iccs based on exclusions

```{r}

icc_long <- iccs %>% 
        gather(key_t, value_t, c(ICC.t, ICC.t_unknown, ICC.t_1st, ICC.t_1stunknown)) %>%
        gather(key_p, value_p, c(ICC.p, ICC.p_unknown, ICC.p_1st, ICC.p_1stunknown))

summary(lm(value_t ~ as.factor(key_t), 
           data = (icc_long %>% 
                           filter(stim == 'profile' & study == 1))))
summary(lm(value_t ~ as.factor(key_t), 
           data = (icc_long %>% 
                           filter(stim == 'profile' & study == 2))))


summary(lm(value_p ~ as.factor(key_p), 
           data = (icc_long %>% 
                           filter(stim == 'profile' & study == 1))))
summary(lm(value_p ~ as.factor(key_p), 
           data = (icc_long %>% 
                           filter(stim == 'profile' & study == 2))))


summary(lm(value_t ~ as.factor(key_t), 
           data = (icc_long %>% 
                           filter(stim == 'followers'))))
### this one makes a difference:
summary(lm(value_p ~ as.factor(key_p), 
           data = (icc_long %>% 
                           filter(stim == 'followers'))))

summary(lm(value_t ~ as.factor(key_t), 
           data = (icc_long %>% 
                           filter(stim == 'friends'))))
summary(lm(value_p ~ as.factor(key_p), 
           data = (icc_long %>% 
                           filter(stim == 'friends'))))

```

Look at correlations among iccs?

Over everything (not especially meaningful, probably):
```{r}
cor(iccs[c(4,6,8,10,5,7,9,11)])
```

By study/stimuli:

```{r}
iccs %>% group_by(study, stim) %>% 
        do({
                data.frame(cor(iccs[c(4,6,8,10,5,7,9,11)]))
                })
```

Calculate differences, to see how big the changes are.
```{r}
iccs <- iccs %>% mutate(diff_allunknown.t = ICC.t - ICC.t_unknown,
                        diff_all1st.t = ICC.t - ICC.t_1st,
                        diff_allunknown1st.t = ICC.t - ICC.t_1stunknown,
                        diff_1stunknown.t = ICC.t_1st - ICC.t_unknown,
                        diff_1stunknown1st.t = ICC.t_1st - ICC.t_1stunknown,
                        diff_unknown1stunknown.t = ICC.t_unknown - ICC.t_1stunknown,
                        diff_allunknown.p = ICC.p - ICC.p_unknown,
                        diff_all1st.p = ICC.p - ICC.p_1st,
                        diff_allunknown1st.p = ICC.p - ICC.p_1stunknown,
                        diff_1stunknown.p = ICC.p_1st - ICC.p_unknown,
                        diff_1stunknown1st.p = ICC.p_1st - ICC.p_1stunknown,
                        diff_unknown1stunknown.p = ICC.p_unknown - ICC.p_1stunknown)


describe(iccs %>% filter(study == 1 & stim == 'profiles'))[1:4]
describe(iccs %>% filter(study == 1 & stim == 'friends'))[1:4]
describe(iccs %>% filter(study == 1 & stim == 'followers'))[1:4]

describe(iccs %>% filter(study == 2))[1:4]
```

